---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document: default
  pdf_document: default
---

# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv",header = T,sep = ",")
summary(activity)
```

## What is mean total number of steps taken per day?

```{r}
act1 <- aggregate(steps ~ date, activity, sum)
?hist
head(act1)
hist(act1$steps, breaks = 30, xlab = "Steps", ylab = "Counts", main = "Number of steps taken" ,col = "yellow")
```

```{r}
## Mean
mean(act1$steps)

## Median
median(act1$steps)
```


## What is the average daily activity pattern?
```{r}
act2 <- aggregate(steps ~ interval , act,mean)
head(act2)
```

## Make a time series plot (i.e.type="l") of the 5-minute interval
## (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
plot(act2$interval, act2$steps,type = "l", xlab = "Intervel", 
     ylab = "Average step taken", main = "Time Series Plot")
     ```

## Which 5-minute interval, on average across all the days in the dataset, 
## contains the maximum number of steps?

act2[act2$steps ==max(act2$steps),]
```

# Imputing missing values

## Calculate and report the total number of missing values in the dataset 
## (i.e. the total number of rows with NAs)
```{r}
nrow(act [is.na(act$steps)==TRUE,])
```

## Devise a strategy for filling in all of the missing values in the dataset.
## The strategy does not need to be sophisticated. For example, you could use 
## the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r}
library(Hmisc)
```
## Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
act_new<- act
act_new$steps <- with(act_new, impute(steps, mean))
sub_act <- aggregate(steps ~ date, act_new, sum)
sub_act
```

## Make a histogram of the total number of steps taken each day and Calculate and 
## report the mean and median total number of steps taken per day. 
## Do these values differ from the estimates from the first part of the assignment? 
## What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
hist(sub_act$steps, breaks = 20,xlab = "Steps",
     ylab = "Counts",main = "New Histogram of Steps after imputing")

mean(sub_act$steps)

median(sub_act$steps)
```

# Are there differences in activity patterns between weekdays and weekends?
```{r}
act$date<-with(act,weekdays(as.Date(act$date)) %in% c('Sunday','Saturday'))
```

#create plot
```{r}
xyplot(steps ~ interval | date, data = act_new, layout = c(1, 2),type = "l",
       strip=strip.custom(factor.levels=c("WEEKEND","WEEKDAY")))
```

